---
title: "Workplan"
author: "Anders G. Finstad"
date: "February 14, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data capture forms

### Rationale / idea

* While DwC and the DwC-A is nice for data publishing, archiving and exchange, it's often not possible to or wanted to punch (a.k.a. enter your data into this format directly).

* Standarized and documented forms (spredsheets) for entering data from the most used ecological sampling technices would eases the process of writing mapping rutines for publishing data

* Forms (spreadsheets) should contain validatation rules to ensure minimum of work afterwards 

* Generators developed in R and implementented in shiny

### Worlflow

#### 1. load libraries and set define input parameters (template etc.)
```{r input_param, echo=TRUE, results=FALSE}
library(openxlsx) # use openxlsx (no nasty java dependencies)
library(stringi)
library(stringr)
library(dplyr)
library(uuid)

# innstillinger
template_file <- "./data/templates/gillnet-surveys/NINA_gillnet_survey_template.csv"
maxN <- 100
template_folder <- "./data/templates/gillnet-surveys/"

## header style
hs <- createStyle(fontColour = "#ffffff", fgFill = "#4F80BD", halign = "center", valign = "center", textDecoration = "Bold", border = "TopBottomLeftRight", textRotation = 90)

```

#### 2. Create workbook, sheets and column names 


```{r create_workbook, echo=TRUE, results=FALSE}

# create workbook and worksheets
template <- read.csv(template_file)
worksheets <- unique(template$sheet)

wb <- createWorkbook()


## add worksheets
for(i in 1:length(worksheets)){
  addWorksheet(wb, paste(worksheets[i]))
  setColWidths(wb, sheet = paste(worksheets[i]), cols=1:100, widths = "auto")
}

## add column names
for(i in 1:length(worksheets)){
  xxx <- matrix(ncol=length(template$field_name_NO[template$sheet==paste(worksheets[i])])) %>%
  as.data.frame()
  names(xxx) <- as.character(template$field_name_NO[template$sheet==paste(worksheets[i])])
  writeDataTable(wb, sheet = i, x = xxx, firstColumn=TRUE)#, headerStyle=hs)
}

```

#### 3. add data-validation, lookup columns etc..

Loops trough the "validation" column of the template sheet

* ID:UUID - generates a set of UUIDs (of length maxN) to be used as GUIDs for the ID, stores the resulting IDs in sepparate lookup sheets
* '*.csv' - looks for files with controlled vocabulary. Using the first column of the table as refference
* must_contain - column must contain value, typically refferenceing ID of highter hiarcical structure 
* unique - value must be unique (typically ID columns)

```{r lookup_etc, echo=TRUE, results=FALSE}


##--------------------------------------
## add lookupdata and data validatation
##--------------------------------------

for(i in 1:length(template$validation)){
  
  ###-------------------------------------------
  ### IDs - add UUID if validation = lookup:UUID
  ###-------------------------------------------
  if(template$validation[i]=="ID:UUID"){
    addWorksheet(wb, paste0("l_",template$field_name_NO[i]))
    writeData(wb, sheet = paste0("l_",template$field_name_NO[i]), 
              x = c(paste(template$field_name_NO[i]),sapply(seq_along(1:maxN), UUIDgenerate))
    )
  }
  

  ###--------------------------------------------------------------------------
  ### lookup tables: add validation from controlled vocabulary attached as file
  ###--------------------------------------------------------------------------
    if(template$validation[i] %in% list.files(template_folder)){
      #### add lookupdata
      sheet_name_i <- str_sub(template$validation[i], end = -5)
      sheet_content <- read.csv(paste0(template_folder,"/",template$validation[i]))
      addWorksheet(wb, sheetName=sheet_name_i)
      writeDataTable(wb, sheet = sheet_name_i, x = sheet_content)
      
      ####-------------------
      #### add datavalidation
      ####-------------------
      template_sheet_i <- template %>%
        filter(sheet == template$sheet[i])
      column <- which(template_sheet_i$validation==template$validation[i])
      
      dataValidation(wb, sheet=paste(template$sheet[i]),
                     col = column,
                     rows = 2:31, 
                     type = "list", value = paste0("'",sheet_name_i,"'","!$A$2:$A$",length(sheet_content[,1]))
      )
    }
  
  

  ###--------------------------------------------------------------------------
  ### data validation: add validation from relational structure
  ###--------------------------------------------------------------------------
    if(str_detect(template$validation[i],"must_contain")){
      #### refferences?
      ref_sheet <- str_split_fixed(str_sub(template$validation[i],14,-2),"\\$",2)[1]
      ref_column <- str_split_fixed(str_sub(template$validation[i],14,-2),"\\$",2)[2]
      reff_sheet_data <- template %>%
        filter(sheet == ref_sheet)
      reff_column_nr <- which(reff_sheet_data[,2]==ref_column)
      
      template_sheet_i <- template %>%
        filter(sheet == template$sheet[i])
      column <- which(template_sheet_i$validation==template$validation[i])
      
      dataValidation(wb, sheet=paste(template$sheet[i]),
                     col = column,
                     rows = 2:maxN, 
                     type = "list", 
                     value = paste0("'",ref_sheet,"'",
                                    "!$",LETTERS[reff_column_nr],
                                    "$2:$"
                                    ,LETTERS[reff_column_nr],
                                    "$"
                                    ,maxN)
      )
            
}
  ###-----------------------------------------------
  ### Refferences and lookup - add refference columns 
  ###-----------------------------------------------
   if(str_detect(template$validation[i],"lookup")){
     
     template_sheet_i <- template %>% filter(sheet == template$sheet[i])
      column <- which(template_sheet_i$validation==template$validation[i])
     
     vlookup_parts <- str_sub(template$validation[i],start=11,end=-3) # pick up formula part
     reffreence_column <- column+as.numeric(str_sub(template$validation[i],start=8,end=9))
     v2 <- paste0("VLOOKUP(", LETTERS[reffreence_column],2:maxN,",",vlookup_parts,",0)")
     class(v2) <- c(class(v2), "formula")
     
     writeData(wb, sheet = paste(template$sheet[i]), x = v2, 
               startCol = column, startRow = 2)
     
     style <- createStyle(fgFill = "#B7B7B7")
     addStyle(wb, sheet = paste(template$sheet[i]), style=style, 
              rows=1:maxN, cols=column, gridExpand = FALSE, stack = FALSE)
     
   }
  
  ###-----------------------------------------------
  ### Unique values  - typically IDs 
  ###-----------------------------------------------
   if(str_detect(template$validation[i],"unique")){
     ## hm..... 
   }

  ###-----------------------------------------------
  ### Is date
  ###-----------------------------------------------
   if(str_detect(template$validation[i],"is_date")){
        template_sheet_i <- template %>%
        filter(sheet == template$sheet[i])
      column <- which(template_sheet_i$validation==template$validation[i])
      
      dataValidation(wb, sheet=paste(template$sheet[i]),
                     col = column,
                     rows = 2:maxN, 
                     type = "date",
                     value = as.Date(Sys.Date()),
                     operator = "lessThan"
      )
   }
  
     if(str_detect(template$validation[i],"is_year")){
        template_sheet_i <- template %>%
        filter(sheet == template$sheet[i])
      column <- which(template_sheet_i$validation==template$validation[i])
      
      dataValidation(wb, sheet=paste(template$sheet[i]),
                     col = column,
                     rows = 2:maxN, 
                     type = "whole",
                     value = c(1000,as.numeric(str_sub(Sys.Date(),start=1,end=4))),
                     operator = "between"
      )
     }
  
       if(str_detect(template$validation[i],"is_month")){
        template_sheet_i <- template %>%
        filter(sheet == template$sheet[i])
      column <- which(template_sheet_i$validation==template$validation[i])
      
      dataValidation(wb, sheet=paste(template$sheet[i]),
                     col = column,
                     rows = 2:maxN, 
                     type = "whole",
                     value = c(1,12),
                     operator = "between"
      )
       }
  
       if(str_detect(template$validation[i],"is_day")){
        template_sheet_i <- template %>%
        filter(sheet == template$sheet[i])
      column <- which(template_sheet_i$validation==template$validation[i])
      
      dataValidation(wb, sheet=paste(template$sheet[i]),
                     col = column,
                     rows = 2:maxN, 
                     type = "whole",
                     value = c(1,31),
                     operator = "between"
      )
   }
  
}

```



```{r save_file, echo=TRUE, results=FALSE}

##--------------------
## Save file 
##---------------------

# random unique code for each workbook download
workbookID <- stri_rand_strings(1,4,pattern="[A-Za-z0-9]") 
workbook_name <- paste0("testform_",workbookID,".xlsx")
workbook_name <- "testform.xlsx"
# save/ dowload workbook


saveWorkbook(wb, workbook_name, overwrite = TRUE)



```

